# 01/07/23

## Plan today

------------------------------------------------------------------------

Verity witihn dataset integration and clustering

------------------------------------------------------------------------

## Verity within integration

commands are running fine, but taking alot of time espdcially at the CCA stage. If I use RPCA instead which is faster, the [documentation](https://satijalab.org/seurat/articles/integration_rpca.html) mentions some conditions "\* A substantial fraction of cells in one dataset have no matching type in the other \* Datasets originate from the same platform (i.e. multiple lanes of 10x genomics) \* There are a large number of datasets or cells to integrate (see INSERT LINK for more tips on integrating large datasets)".

Alterations to the code would include this after defining features:

```{r}
#Perform PCA on the data individually before integration
Ver_list <- lapply(X = Ver_list, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})

#USe RPCA
Verity.anchors <- FindIntegrationAnchors(object.list = Ver_list, anchor.features = features, verbose=TRUE, reduction='rpca')
```

(Original)

```{r}
library(DropletUtils)
library(Seurat)
library(tidyverse)
library(dplyr)
library(stringr)
library(knitr)
library(DoubletFinder)
library(SoupX)
library(scCustomize)
```

### Make metadata for treatment

```{r}
sampleLabs<-names(Ver_singlets_merged@active.ident)

#FIXME change these to be treatment levels of experimental setup

#Assigning names to the samples to keep identity
#From Chat GPT
treat_detect <- case_when(
  str_detect(sampleLabs, "sample11") ~ "AGEING",
  str_detect(sampleLabs, "sample1") ~ "NBH",
  str_detect(sampleLabs, "sample2") ~ "NBH",
  str_detect(sampleLabs, "sample5") ~ "ME7",
  str_detect(sampleLabs, "sample6") ~ "ME7",
  str_detect(sampleLabs, "sample9") ~ "AGEING",
  TRUE ~ NA_character_
)
#
Ver_singlets_merged@meta.data$treatment<-treat_detect

#FIXME may not be needed if the integration workflow specifies what to consider anyway
Idents(object=Ver_singlets_merged)<-"treatment"
rm(sampleLabs, treat_detect)
```

## Within dataset clustering

```{r}
# adapted from https://satijalab.org/seurat/articles/integration_introduction.html

#FIXME set the RNA assay as active
DefaultAssay(Ver_singlets_merged) <- "RNA"

# split the dataset into a list of three seurat objects (ME7, NBH, AGEING)
Ver_list<-SplitObject(Ver_singlets_merged, split.by = "treatment")
rm(Ver_singlets_merged)

# normalize and identify variable features for each dataset independently
Ver_list <- lapply(X = Ver_list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = Ver_list)

#identify anchors
Verity.anchors <- FindIntegrationAnchors(object.list = Ver_list, anchor.features = features, verbose=TRUE)

# this command creates an 'integrated' data assay using the anchors
Verity.combined <- IntegrateData(anchorset = Verity.anchors)

#save the object
save(Verity.combined,"Rdata/Verity.self_integrated.Rdata")

#Perform integrated analysis
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(Verity.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
Verity.combined <- ScaleData(Verity.combined, verbose = FALSE)
Verity.combined <- RunPCA(Verity.combined, npcs = 30, verbose = FALSE)
Verity.combined <- RunUMAP(Verity.combined, reduction = "pca", dims = 1:30)
Verity.combined <- FindNeighbors(Verity.combined, reduction = "pca", dims = 1:30)
Verity.combined <- FindClusters(Verity.combined, resolution = 0.5)

# Visualization
p1 <- DimPlot(Verity.combined, reduction = "umap", group.by = "treatment")
p2 <- DimPlot(Verity.combined, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2

#Visualise treatments side by side
DimPlot(Verity.combined, reduction = "umap", split.by = "treatment")
```

## SUMMARY
